---
title: "Example Interactive Graphs"
author: "Matthew Martinez"
date: "03/30/2021"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packs <- c("googleVis", "tidyr", "gridExtra", "dplyr", "plyr", "stringr", "data.table", "ggplot2", "scales", "plotly", "gganimate")
for ( p in packs) library(p, character.only = T)
rm(packs, p)

```

### Plot_ly and GoogleVis
In this lesson we will be learning how to visualize data using Plot_ly and GoogleVis. We will do this by looking at some built-in examples in R and by looking at how I have incorporated these in a project I've done at work. Keep in mind that for the example below I have made up the data, but the overall structure of the report is kept in tact.

If we have time we will also touch on mapping in Leaflet. If not you can find the [examples here](https://rpubs.com/uky994/594590).

## **Transitions Analysis**

I can discuss what this study is about **here**. Including formatting and working with R Markdown within this example. [Click here for RStudio](www.rstudio.com)

Next, I discuss the results of this assessment.


```{r data_clean, echo=FALSE, include=FALSE}

#Bring in Data
all_cohorts<-read.csv("https://raw.githubusercontent.com/mattdemography/STA_6233_Spring2021/master/Data/all_cohorts.csv")
count<-read.csv("https://raw.githubusercontent.com/mattdemography/STA_6233_Spring2021/master/Data/cohort_counts.csv")

#Find All Possible Transitions

caps<-names(all_cohorts[1:13])
caps_l<-names(all_cohorts[2:14]) #Lag Previous Variable
j<-as.data.frame(table(all_cohorts$trans))

#Label these transitions
  #Create list to change to
  cap_num<-c("13", "15", "19", "21", "25", "31", "33", "37", "39", "42", "46", "47", "48", "49", "51", "73", "3", "4", "5", "6", "7")
  cap_name<-c("Non-Exempt ->", "Part-Time ->",  "Ex-Employee ->", "Unknown ->",  "Temp-Employee ->", "Temp-Employee ->", "Intern ->", "Intern ->", "Unknown ->", 
              "Deceased-Employee ->", "Ex-Employee ->", "Outsourced ->", "Temp ->", "Ex-Temp ->", "Business-Partner ->", "Deceased-Temp ->", 
              "Full-Time ->", "Unknown ->", "Ex-Intern ->", "Family ->", "Ex-Employee ->")
#Create Loop
  j$name<-j$Var1 #Create Copy
  for(s in 1:length(cap_num)){
    j$name<-gsub(cap_num[s], cap_name[s], j$name) #Use Regular Expression to take what is in cap_num and put cap_name instead
  }
  #Eliminate all same-state non-transition months
  j$unique_name<-vapply(strsplit(j$name, " "), function(x) paste(unique(x), collapse = " "), character(1L)) 
  
#Get All Unique Transitions overall and how often it occurs
  t<-aggregate(j$Freq, by=list(j$unique_name), FUN=sum)
  t<-plyr::rename(t, c("Group.1"="Transition Type", "x"="Frequency"))
  t<-t[order(-t$Frequency),]

#Add a total Column to count
count$Total<-rowSums(count[2:13])

#Create a Survival Format
c<-count$Total
  #Create a How many part-time employees remain after end of month
  for(i in 1:1){
    eval(parse(text=paste0('count$ms_', i-1, '<-c')))
    for(k in 1:12){
      eval(parse(text=paste0('count$ms_', k, '<-(count$ms_', k-1, ' - count$Month_', k, ')')))
    }
  }
  #What Percent of total that became full-time at end of month
  for(k in 1:12){
    eval(parse(text=paste0('count$Percent_Left_in_Month_', k,'<-round(as.numeric(count$Month_', k,'/count$Total)*100,2)')))
  }
  #What percent remain as non full-time at end of month
  for(k in 1:12){
    eval(parse(text=paste0('count$Percent_Remain_in_Month_', k,'<-round(as.numeric(count$ms_', k-1,'/count$Total)*100,2)')))
  }

#Now Make this data long
  tout<- count[2:15] %>% gather(Month_Code, No_Longer_Employee, Month_1:Month_12)
  tout_p<- count[29:40] %>% gather(Month_Code, Percent_No_Longer_Employee, Percent_Left_in_Month_1:Percent_Left_in_Month_12)
  survive<-count[14:27] %>% gather(time_survive, Still_Employee, ms_0:ms_11)
  survive_p<-count[41:52] %>% gather(time_survive, Percent_Still_Employee, Percent_Remain_in_Month_1:Percent_Remain_in_Month_12)
  count_l<-cbind(tout[c(1:2,4)], tout_p[2], survive[4], survive_p[2])
  count_l$Month<-with(count_l, ave(as.character(count_l$cohort), as.character(count_l$cohort), FUN=seq_along)) %>% as.numeric()
  count_l$Month<-count_l$Month + 100 #Adding 100 so that it remains in order as 1, 11, 12, 2 will be order without this

  all_sum1<-mean(count$Total)
  all_sum2<-mean(all_cohorts$count)
  tot_cnt<-sum(count$Total)
  
  #Reorder Factor Levels
  count$fl<-ifelse(count$cohort=="17-Jan", 1, ifelse(count$cohort=="17-Feb", 2, ifelse(count$cohort=="17-Mar", 3,
    ifelse(count$cohort=="17-Apr", 4, ifelse(count$cohort=="17-May", 5, ifelse(count$cohort=="17-Jun", 6, ifelse(count$cohort=="17-Jul", 7,
    ifelse(count$cohort=="17-Aug", 8, ifelse(count$cohort=="17-Sep", 9, ifelse(count$cohort=="17-Oct", 10, ifelse(count$cohort=="17-Nov", 11,
    ifelse(count$cohort=="17-Dec", 12, NA))))))))))))
  count$cohort<-factor(count$cohort, levels=count$cohort[order(count$fl)])
  
  count_l$fl<-ifelse(count_l$cohort=="17-Jan", 1, ifelse(count_l$cohort=="17-Feb", 2, ifelse(count_l$cohort=="17-Mar", 3,
    ifelse(count_l$cohort=="17-Apr", 4, ifelse(count_l$cohort=="17-May", 5, ifelse(count_l$cohort=="17-Jun", 6, ifelse(count_l$cohort=="17-Jul", 7,
    ifelse(count_l$cohort=="17-Aug", 8, ifelse(count_l$cohort=="17-Sep", 9, ifelse(count_l$cohort=="17-Oct", 10, ifelse(count_l$cohort=="17-Nov", 11,
    ifelse(count_l$cohort=="17-Dec", 12, NA))))))))))))

#Grab Only the 6th month
  six<-count_l[61:72,c(1,6)]
  max<-subset(six, six$Percent_Still_Employee==max(six$Percent_Still_Employee))
  min<-subset(six, six$Percent_Still_Employee==min(six$Percent_Still_Employee))
  
```

### Results

Figure #1 below displays the totals by month of newly added people. In total for 2017, there were `r as.numeric(tot_cnt)` people analyzed across 12 cohorts. The average across cohorts is `r round(all_sum1, 2)` and is displayed as a red dashed line in the figure below. The largest cohort is March17 at 3,291 new people and the smallest is November17 at 2,143. Any cohort that is above the average is displayed in darker blue and cohorts below average are displayed in lighter blue.

```{r visuals_1, echo=FALSE, include=TRUE, warning=FALSE}
all_sum1<-mean(count$Total)
bar1<-ggplot(data=count, aes(count$cohort, fill=count$cohort)) + geom_bar(aes(weight=count$Total)) +  
  scale_fill_manual(values=c("17-Apr"="lightcyan3", "17-Aug"="lightcyan4", "17-Dec"="lightcyan3", "17-Feb"="lightcyan4", "17-Jan"="lightcyan3",
                             "17-Jul"="lightcyan4", "17-Jun"="lightcyan3", "17-Mar"="lightcyan4", "17-May"="lightcyan3", "17-Nov"="lightcyan4",
                             "17-Oct"="lightcyan3", "17-Sep"="lightcyan4"), guide=F) + xlab("Cohort") + ylab("Total Employee") + theme_light() + 
  scale_y_continuous(limits=c(0, 3500), oob=rescale_none) + geom_hline(yintercept = all_sum1, linetype="dashed", colour="indianred3") + 
  geom_text(aes(x=count$cohort, y=count$Total, label=count$Total), vjust=-1) + ggtitle("Total Number of Employees as Operators by Cohort")

ggplotly(bar1)
```

Figure #2 displays the average months that people stay as employees (as opposed to changing statuses) by cohort. The average across all 12 cohorts is `r round(all_sum2, 2)` and is displayed as a red dashed line in Figure #2. The highest average number of months that employees stayed in their current roles as 8.99 and occurred in the November17 and September17 cohorts. As above, any cohort that is below the global mean is displayed in light blue while any cohort above the mean is shown in dark blue.

```{r visuals_2, echo=FALSE, include=TRUE, warning=FALSE}

  #Create new dataset with just the summary statistics
sum_stats<-all_cohorts[,18:19] %>% group_by(cohort) %>% summarise_each(funs(mean(., na.rm = T), sd(., na.rm=T)))
sum_stats$fl<-ifelse(sum_stats$cohort=="17-Jan", 1, ifelse(sum_stats$cohort=="17-Feb", 2, ifelse(sum_stats$cohort=="17-Mar", 3,
  ifelse(sum_stats$cohort=="17-Apr", 4, ifelse(sum_stats$cohort=="17-May", 5, ifelse(sum_stats$cohort=="17-Jun", 6, ifelse(sum_stats$cohort=="17-Jul", 7,
  ifelse(sum_stats$cohort=="17-Aug", 8, ifelse(sum_stats$cohort=="17-Sep", 9, ifelse(sum_stats$cohort=="17-Oct", 10, ifelse(sum_stats$cohort=="17-Nov", 11,
  ifelse(sum_stats$cohort=="17-Dec", 12, NA))))))))))))
sum_stats$cohort<-factor(sum_stats$cohort, levels=sum_stats$cohort[order(sum_stats$fl)])
#  all_sum2<-mean(all_cohorts$count)
  
all_sum2<-mean(all_cohorts$count)
bar2<-ggplot(data=sum_stats, aes(sum_stats$cohort, fill=sum_stats$cohort)) + geom_bar(aes(weight=sum_stats$mean)) +  
  scale_fill_manual(values=c("17-Apr"="lightcyan3", "17-Aug"="lightcyan4", "17-Dec"="lightcyan3", "17-Feb"="lightcyan4", "17-Jan"="lightcyan3",
                             "17-Jul"="lightcyan4", "17-Jun"="lightcyan3", "17-Mar"="lightcyan4", "17-May"="lightcyan3", "17-Nov"="lightcyan4",
                             "17-Oct"="lightcyan3", "17-Sep"="lightcyan4"), guide=F) + xlab("Cohort") + ylab("Average Months") + theme_light() + 
  scale_y_continuous(limits=c(8.25,9.25), oob=rescale_none) + geom_hline(yintercept = all_sum2, linetype="dashed", colour="indianred3") + 
  geom_text(aes(x=sum_stats$cohort, y=sum_stats$mean, label=round(sum_stats$mean,2)), vjust=-1) + ggtitle("Average Months Employee Remain by Cohort") 
plot(bar2)
```

Next we see Figure #3 which displays the percent of employees that remain over time. The month number on the x-axis corresponds to the month of observation for each cohort. For example, month number 102 is the second month for the Feb17 cohort which would be March17, but the second month for the July17 cohort is August17 and so on. This allows us to view trends over time per cohort. The dashed vertical line symbolizes the 6th month, or when employees should convert to full-time status according to policy rules. As one can see for all cohorts, at least 70% of employees remain as part-time after 6 months. The `r max[,1]` cohort has the largest percentage of part-time employees remaining at `r max[,2]` and the `r min[,1]` cohort has the least part-time employees remaining at `r min[,2]`.

```{r visuals_3, echo=FALSE, warning=FALSE}

  line_t<-ggplot(data=count_l, aes(x=count_l$Month, y=count_l$Percent_Still_Employee, group=count_l$cohort)) + xlab("Month Number") +
    ylab("Percent Still Employee") +  geom_line(aes(color=count_l$cohort)) + geom_vline(xintercept = 106, linetype="dashed") + theme_light() + 
    scale_color_discrete(name="Legend") + scale_x_continuous(breaks=seq(101, 112,1)) + ggtitle("Figure #3 Percent Employees Remaining After Each Period by Cohort")
  plot(line_t)
```

I am providing Google Visualizations. Unfortunately after creation of this lesson Adobe Flash has be deprecated and this has affected Google MotionChart. MotionChart was a unique way to visualize data movement over time. Instead we will be using plot_ly which can accomplish the same task, just with a bit more work in coding. Let's build it together in class!

```{r visuals_4, echo=FALSE, warning=FALSE}
#Create Google Visualizations  
  #m_reg<-gvisMotionChart(count_l[1:132,c(1:8)], idvar="cohort", timevar = "Month")
  #m_time<-gvisMotionChart(count_l[,c(1:2, 5:7)], idvar="cohort", timevar = "Month")
  #plot(m_reg)
  #plot(m_time)
  
  p <- ggplot(count_l, aes(x = Still_Employee, y=No_Longer_Employee, colour=cohort)
              )  +  geom_point(show.legend = FALSE, alpha = 0.7)
  p

  ex<-count_l[13:144,]
  plot_ly(x = ex$Month, y = ex$Percent_Still_Employee, color = ex$cohort, type = 'scatter', mode = 'lines')
 
#plot_ly(x = stocks$time, y = stocks$price, color = stocks$index, type = 'scatter', mode = 'lines')


```

### Appendix

I am providing the following table that displays the top 15 transitions made from the part-time status. As can be seen, the data contain some missing values ('NA'). This can potentially be an important factor when analyzing the data above. These NA's can mean one of two things, each with different affects on the analyses. The NAs could mean that the previous status is still in effect or they could mean that there was a transition but this status change was not correctly documented.

```{r visuals_5, echo=FALSE, warning=FALSE}
  myTheme<-ttheme_default(
    core=list(fg_params=list(hjust=1, x=1),
              bg_params=list(fill=c("cornflowerblue", "gray92"))),
    colhead=list(fg_params=list(col="white"),
                 bg_params=list(fill="mediumpurple"))
  )
  
  trans_table<-tableGrob(t[1:15,], rows=NULL, theme=myTheme)
  trans_table<-plot_ly(type="table",header=list(values=names(t)), cells=list(values=unname(t[1:15,]))) 

  trans_table
```